---
title: Assessment 3 - Case Study - Super Netball 2020 
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
---

```{r setup, include=FALSE}
# Firstly ensure all required packages are loaded in the library
library(tidyverse)
library(ggplot2)
library(dplyr)
library(readr)
library(viridis)
library(DT)
library(plotly)
library(leaflet)
library(flexdashboard)

# Load up the data set 
SuperNetball_data <- read_csv("Dataset3_Assessment3.csv", show_col_types = FALSE)

# Visualise the entire Super Netball data set within the source to ensure it is loaded
datatable(SuperNetball_data)

# Clean the Super Netball data set and see if their are any missing values
sum(is.na(SuperNetball_data))
# No missing data found 

# Convert both 'Athlete' and 'Team' to categorical variables this allows my R analysis to become more efficient, accurate, and visually clear.  It correctly groups data to assist in visualisations.
SuperNetball_data <- SuperNetball_data %>%
  mutate(Athlete = as.factor(Athlete),
         Team = as.factor(Team),
         Statistic = as.factor(Statistic))

# Create Team Performance data set 
team_performance <- SuperNetball_data %>%
  group_by(Team) %>%
  summarise(Total_Performance = sum(Total, na.rm = TRUE)) %>%
  arrange(desc(Total_Performance))

# Ensure no duplicate rows
SuperNetball_data <- distinct(SuperNetball_data)

# Using str function ensure the structure of the dataset is correct
str(SuperNetball_data)

#------------------------------------------------------------------------------#
# Look at the Statisical Distribution of the Super Netball data by creating a boxplot
#------------------------------------------------------------------------------#
# Create a Boxplot of Statistical Distribution and then develop this using colourblind friendly colours.
# This boxplot will help with essential understanding of the distribution of player performance in Netball
# It will help the coach identify strengths, weaknesses and anomalies
# It can help to enhance decision making for player development, game startegy and recruitment of players.

# Display the entire dataset using datatable 
datatable(SuperNetball_data, options = list(scrollX = TRUE, pageLength = 10))
```

Column {data-width=650}
-----------------------------------------------------------------------

### Chart A
ggplot(SuperNetball_data, aes(x = Statistic, y = Total, fill = Statistic)) +
  geom_boxplot(alpha = 0.8) +
  scale_fill_viridis_d(option = "D", begin = 0.1, end = 0.9) +
  labs(title = "Boxplot of Statistical Distribution",
       x = "Statistic",
       y = "Total Values") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 10),
    axis.title = element_text(size = 14, face = "bold"),
    plot.title = element_text(size = 16, face = "bold", hjust = 0.5),
    legend.position = "none"
  )

```{r}

#------------------------------------------------------------------------------#
# Summarise the data to provide the coach with some easy essential data to help with planning based on important statistics 
#------------------------------------------------------------------------------#
# Discover using code the Top 10 players and teams based on total performance 
# Using the filter and pivot wider commands summarise the shooting performance of Super Netball 2020 season

# Convert data into wide format so that each athlete has separate shooting statistics

# Define shooting-related statistics
scoring_stats <- c("goal_from_zone1", "goal_from_zone2", "attempt_from_zone1", "attempt_from_zone2")

# Filter only shooting-related statistics
shooting_data <- SuperNetball_data %>%
  filter(Statistic %in% scoring_stats)

# Convert data into wide format so that each athlete has separate shooting statistics
shooting_summary <- shooting_data %>%
  pivot_wider(names_from = Statistic, values_from = Total, values_fill = list(Total = 0))

# Ensure required columns exist before using them
required_columns <- c("goal_from_zone1", "goal_from_zone2", "attempt_from_zone1", "attempt_from_zone2")

shooting_summary <- shooting_summary %>%
  mutate(
    goal_from_zone1 = coalesce(goal_from_zone1, 0),
    goal_from_zone2 = coalesce(goal_from_zone2, 0),
    attempt_from_zone1 = coalesce(attempt_from_zone1, 0),
    attempt_from_zone2 = coalesce(attempt_from_zone2, 0)
  )

# Calculate Shooting Accuracy
shooting_summary <- shooting_summary %>%
  mutate(
    Total_Goals = goal_from_zone1 + goal_from_zone2,
    Total_Attempts = attempt_from_zone1 + attempt_from_zone2,
    Shooting_Accuracy = ifelse(Total_Attempts > 0, (Total_Goals / Total_Attempts) * 100, 0)
  )

# Identify the Top 10 Shooters by Accuracy - create top shooters data 
top_shooters <- shooting_summary %>%
  arrange(desc(Shooting_Accuracy)) %>%
  slice_head(n = 10)

# Display Top 10 Shooters
view(top_shooters)

# Look at general play and performance both team and players 
# Summary of the 'Total' column
summary(SuperNetball_data$Total)

# Ensure no missing columns in team performance
team_performance <- SuperNetball_data %>%
  group_by(Team) %>%
  summarise(Total_Performance = sum(Total, na.rm = TRUE)) %>%
  arrange(desc(Total_Performance))

# check column names 
print(colnames(team_performance))

# Top 10 Players Based on Total Performance
top_players <- SuperNetball_data %>%
  group_by(Athlete) %>%
  summarise(Total_Sum = sum(Total, na.rm = TRUE)) %>%
  arrange(desc(Total_Sum)) %>%
  slice_head(n = 10)

# Display Top 10 Athletes
view(top_players)

# Create a table sumarrising team perforamnce 
datatable(team_performance)

#Create a data table of top performing players 
datatable(top_players)

# Create the Top Performing Athletes Data 
# Ensure no errors if top_players is empty
if (nrow(top_players) > 0)
  top_players_name <- top_players$Athlete[1]
  
  # Filter Data for the Top Athlete
  top_player_data <- SuperNetball_data %>%
    filter(Athlete == top_players_name)
  
  print(paste("Performance Data for:", top_players_name))
  view(top_player_data)


# Create and cinsider the most recorded statistics 
most_recorded_statistics <- SuperNetball_data %>%
  group_by(Statistic) %>%
  summarise(Total_Count = sum(Total, na.rm = TRUE)) %>%
  arrange(desc(Total_Count)) %>%
  slice_head(n = 10)

# Display Most Recorded Statistics
view(most_recorded_statistics)

datatable(most_recorded_statistics)
```

Column {data-width=350}
-----------------------------------------------------------------------

### Chart B
#------------------------------------------------------------------------------#
# Create a simple Bar Chart as a visualisation for the top 10 players by performance 
#------------------------------------------------------------------------------#
ggplot(top_players, aes(x = reorder(Athlete, -Total_Sum), y = Total_Sum, fill = Athlete)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  scale_fill_viridis_d(option = "C") +  # Colorblind-friendly
  labs(title = "Top 10 Players by Performance",
       x = "Athlete",
       y = "Total Score") +
  theme_minimal()

```{r}

#------------------------------------------------------------------------------#
# Create a simple Bar Chart as a visualisation for the top 10 teams by performance 
#------------------------------------------------------------------------------#
# The bar chart simplifies complex data into an easy to read format for coaches 
# Allows the coaches to quickly identify strongest and weakest teams.
# This may help with coaching, strategy and decision making
# Provides data to back opinion and can be a visualisation for the team 
# Add coloublind friendly colours and hide the legend since x axis already has the label for teams 

```

### Chart C
ggplot(team_performance, aes(x = reorder(Team, Total_Performance), y = Total_Performance, fill = Team)) +
  geom_bar(stat = "identity", show.legend = FALSE) +  
  coord_flip() +
  scale_fill_viridis_d(option = "C") +
  labs(title = "Total Performance by Team", x = "Team", y = "Total Score") +
  theme_classic() +
  theme(
    axis.text.x = element_text(size = 12, face = "bold"),
    axis.text.y = element_text(size = 12, face = "bold"),
    axis.title = element_text(size = 14, face = "bold"),
    plot.title = element_text(size = 16, face = "bold", hjust = 0.5)
  )


```{r}
#------------------------------------------------------------------------------#
# Develop Performance Analytics to assist Netball Coaches in important decision making
#------------------------------------------------------------------------------#

# ---------------------------------------------------------------------------- #
# 1. Shooting Efficiency & Accuracy Analysis
# ---------------------------------------------------------------------------- #
# Looking at Shooting Efficiency and Accuracy
# Coaches would use this information to identify the most efficient and accurate shooters which may be used for Scouting or Recruitment purposes.

# Now we will look at statistics including goals and attempts from Zone 1 and Zone 2 .  This will show us who us efficient vs volume shooters. It will also distinguish between close range and long range threats and enable accurate calculation of shooting percentages  
# This may help the coach design better offensive strategies and is clearly helpful for scouting, recrutiment and player development.

# Statistics including goals & attempts from Zone 1 & Zone 2
scoring_stats <- c("goal_from_zone1", "goal_from_zone2", "attempt_from_zone1", "attempt_from_zone2")

# Using the filter and pivot wider commands summarise the shooting performance of Super Netball 2020 season
shooting_summary <- shooting_summary %>%
  mutate(
    goal_from_zone1 = coalesce(goal_from_zone1, 0),
    goal_from_zone2 = coalesce(goal_from_zone2, 0),
    attempt_from_zone1 = coalesce(attempt_from_zone1, 0),
    attempt_from_zone2 = coalesce(attempt_from_zone2, 0),
    Total_Goals = goal_from_zone1 + goal_from_zone2,
    Total_Attempts = attempt_from_zone1 + attempt_from_zone2,
    Shooting_Accuracy = ifelse(Total_Attempts > 0, (Total_Goals / Total_Attempts) * 100, 0)
  )

if (!"goal_from_zone1" %in% colnames(shooting_summary)) shooting_summary$goal_from_zone1 <- 0
if (!"goal_from_zone2" %in% colnames(shooting_summary)) shooting_summary$goal_from_zone2 <- 0
if (!"attempt_from_zone1" %in% colnames(shooting_summary)) shooting_summary$attempt_from_zone1 <- 0
if (!"attempt_from_zone2" %in% colnames(shooting_summary)) shooting_summary$attempt_from_zone2 <- 0

# Calculate Total Goals, Total Attempts, and Shooting Accuracy
shooting_summary <- shooting_summary %>%
  mutate(
    Total_Goals = goal_from_zone1 + goal_from_zone2,
    Total_Attempts = attempt_from_zone1 + attempt_from_zone2,
    Shooting_Accuracy = ifelse(Total_Attempts > 0, (Total_Goals / Total_Attempts) * 100, 0)
  )

shooting_summary <- shooting_summary %>%
  mutate(
    Total_Goals = goal_from_zone1 + goal_from_zone2,
    Total_Attempts = attempt_from_zone1 + attempt_from_zone2,
    Shooting_Accuracy = ifelse(Total_Attempts > 0, (Total_Goals / Total_Attempts) * 100, 0)
  )

shooting_summary <- shooting_summary %>%
  mutate(Shooting_Accuracy = as.numeric(Shooting_Accuracy))

shooting_accuracy <- shooting_summary %>%
  mutate(
    Total_Goals = goal_from_zone1 + goal_from_zone2,
    Total_Attempts = attempt_from_zone1 + attempt_from_zone2,
    Shooting_Accuracy = ifelse(Total_Attempts > 0, (Total_Goals / Total_Attempts) * 100, 0)
  )

# Select Top 10 Shooters
top_shooters <- shooting_accuracy %>%
  arrange(desc(Shooting_Accuracy)) %>%
  head(10)

# Display Top 10 Shooters
print(top_shooters)
view(top_shooters)

# Create a data table of the op shootere
datatable(top_shooters)

# Display shooting summary viewing the full dataset and then the top 10 shooters 
View(shooting_summary)  
View(head(shooting_summary, 10))  

# ---------------------------------------------------------------------------- #
# 1a. Basic Bar Chart for Top 10 Shooters by Shooting Accuracy
# ---------------------------------------------------------------------------- #
# Bar Chart: Top 10 Shooters by Accuracy
# Add colourblind friendly colours 

# Enhance this plot by adding where the shooting is coming from with a stacked bar plot 
# Standardise athlete names (removes spaces, converts to title case)
# Prevents duplicate athlete entries due to name formatting issues.
# Ensures accurate grouping, filtering, and summarization of player statistics.
# Fixes issues when merging multiple datasets with inconsistent name formats.
# Improves searchability and filtering operations in R.
# Enhances presentation quality for reports, dashboards, and visualizations.
# Standardizing athlete names ensures accuracy, consistency, and professionalism in Super Netball data analysis!```

```

Column {data-width=650}
-----------------------------------------------------------------------

### Chart D
ggplot(top_shooters, aes(x = reorder(Athlete, Shooting_Accuracy), y = Shooting_Accuracy, fill = Athlete)) +
  geom_bar(stat = "identity", show.legend = FALSE) +  
  coord_flip() +
  scale_fill_viridis_d(option = "C") +
  labs(title = "Top 10 Shooters by Accuracy",
       x = "Athlete",
       y = "Shooting Accuracy (%)") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(size = 12, face = "bold"),
    axis.text.y = element_text(size = 12, face = "bold"),
    axis.title = element_text(size = 14, face = "bold"),
    plot.title = element_text(size = 16, face = "bold", hjust = 0.5)
  )


```{r}
# Filter Shooting-Related Statistics
shooting_stats <- c("goal_from_zone1", "goal_from_zone2")

# Create shooting data, data set for ease of summary stats 
shooting_data <- SuperNetball_data %>%
  filter(Statistic %in% shooting_stats)

# Convert Data into Long Format for Stacked Bar Chart which is essential when stacking  multiple categories.  This method allows data readability, simplifies analysis, filtering, grouping and summaries
top_shooters_long <- top_shooters %>%
  pivot_longer(cols = c(goal_from_zone1, goal_from_zone2), names_to = "Zone", values_to = "Goals")

# Create a stacked bar chart that shows the top 10 shooters and showing where the shooting is coming from either zone 1 and zone 2
# Add colour blind friendly colours for readability 


```

Column {data-width=350}
-----------------------------------------------------------------------

### Chart E
ggplot(top_shooters_long, aes(x = reorder(Athlete, Total_Goals), y = Goals, fill = Zone)) +
  geom_bar(stat = "identity", position = "stack", width = 0.8) +  
  coord_flip() +
  scale_fill_viridis_d(option = "C", begin = 0.2, end = 0.9, labels = c("Zone 1 Goals", "Zone 2 Goals")) +  
  labs(title = "Top 10 Shooters: Goals Breakdown by Zone",
       x = "Athlete",
       y = "Total Goals",
       fill = "Scoring Zone") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(size = 12, face = "bold"),
    axis.text.y = element_text(size = 12, face = "bold"),
    axis.title = element_text(size = 14, face = "bold"),
    plot.title = element_text(size = 16, face = "bold", hjust = 0.5),
    legend.title = element_text(size = 12, face = "bold"),
    legend.text = element_text(size = 10)
  )

```{r}

# Filter and extract shooting related statistics goals and attempts from both zone 1 and 2.  This ensures the focus is on shooting which is essential for calculating shooting accuracy.  Allows for comparisaon of close vs long range shooting and create visualisations for scouting and coahcing. 

# Define Shooting Statistics including goals & attempts from Zone 1 & Zone 2
scoring_stats <- c("goal_from_zone1", "goal_from_zone2", "attempt_from_zone1", "attempt_from_zone2")

# Filter Shooting Data
shooting_data <- SuperNetball_data %>%
  filter(Statistic %in% scoring_stats)

# Pivot the Data (ensure missing columns default to 0)
shooting_summary <- shooting_data %>%
  pivot_wider(names_from = Statistic, values_from = Total, values_fill = list(Total = 0)) 

# Check column names to verify if they exist
print(colnames(shooting_summary))

# If any of the expected columns are missing, manually create them with 0
if (!"goal_from_zone1" %in% colnames(shooting_summary)) {
  shooting_summary$goal_from_zone1 <- 0
}
if (!"goal_from_zone2" %in% colnames(shooting_summary)) {
  shooting_summary$goal_from_zone2 <- 0
}
if (!"attempt_from_zone1" %in% colnames(shooting_summary)) {
  shooting_summary$attempt_from_zone1 <- 0
}
if (!"attempt_from_zone2" %in% colnames(shooting_summary)) {
  shooting_summary$attempt_from_zone2 <- 0
}

# Display the first few rows
head(shooting_summary)

shooting_stats <- c("goal_from_zone1", "goal_from_zone2", "attempt_from_zone1", "attempt_from_zone2")

shooting_data <- SuperNetball_data %>%
  filter(Statistic %in% shooting_stats)

# Define shooting-related statistics
scoring_stats <- c("goal_from_zone1", "goal_from_zone2", "attempt_from_zone1", "attempt_from_zone2")

# Check column names to verify `Total_Attempts` exists
print(colnames(shooting_summary))

# Print top shooters to check structure
print(top_shooters)

# ---------------------------------------------------------------------------- #
# 1b. Create a scatter plot of shooting accuracy vs total attempts
# ---------------------------------------------------------------------------- #
# Scatter Plot: Shooting Accuracy vs. Total Attempts
# This scatter pot is designed to show the relationship of how attempts impact shooting accuracy.  This should help coaches indentify not only effiecient shooters but high volume ones as well.
# High volume plus high accuracy would be valued as a high recruit, High Volume but low accuracy potenially inefficient shooter 
# I discovered a way to include the name in the chart to easily identify athelte```


```

### Chart F
ggplot(top_shooters, aes(x = reorder(Athlete, Shooting_Accuracy), y = Shooting_Accuracy, fill = Athlete)) +
  geom_bar(stat = "identity", show.legend = FALSE) +  
  coord_flip() +
  scale_fill_viridis_d(option = "C") +
  labs(title = "Top 10 Shooters by Accuracy",
       x = "Athlete",
       y = "Shooting Accuracy (%)") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(size = 12, face = "bold"),
    axis.text.y = element_text(size = 12, face = "bold"),
    axis.title = element_text(size = 14, face = "bold"),
    plot.title = element_text(size = 16, face = "bold", hjust = 0.5)
  )

```{r}
```

Column {data-width=650}
-----------------------------------------------------------------------

### Chart G
ggplot(top_shooters, aes(x = Total_Attempts, y = Shooting_Accuracy)) +
  geom_point(aes(color = Athlete), size = 4, alpha = 0.9) +  # Scatter plot with colors
  geom_smooth(method = "lm", color = "#D55E00", linetype = "dashed", se = FALSE) +  # Regression line
  geom_text(aes(label = Athlete), vjust = -0.8, hjust = 0.5, size = 3, color = "black") +  # Simple text labels
  scale_color_viridis_d(option = "C", begin = 0.2, end = 0.9) +  # Colorblind-friendly colors
  labs(title = "Shooting Accuracy vs. Total Attempts (Top 10 Shooters)",
       x = "Total Shot Attempts",
       y = "Shooting Accuracy (%)",
       color = "Athlete") +  # Legend for colors
  theme_minimal() +
  theme(
    axis.text.x = element_text(size = 12, face = "bold"),
    axis.text.y = element_text(size = 12, face = "bold"),
    axis.title = element_text(size = 14, face = "bold"),
    plot.title = element_text(size = 16, face = "bold", hjust = 0.5),
    legend.position = "none"  # Hide legend since names are in labels
  )

```{r}
# ---------------------------------------------------------------------------- #
# 2. Turnovers and Possession - Create a turnovers and possession control plot for coaches to visualise
# ---------------------------------------------------------------------------- #
# Create a summary of the turnover related statistics which include bad passes, bad hands, general play turnovers and missed goal turnovers. This will allow coaches to identify teams that may be strugglign with the ball.
turnover_data <- SuperNetball_data %>%
  filter(Statistic %in% c("badPasses", "badHands", "generalPlayTurnovers", "missedGoalTurnover")) %>%
  group_by(Team) %>%
  summarise(Total_Turnovers = sum(Total))

# Display Turnovers by Team
View(turnover_data)
  
# Provide a summary of turnovers by the type and team 
turnover_breakdown <- SuperNetball_data %>%
  filter(Statistic %in% c("badPasses", "badHands", "generalPlayTurnovers", "missedGoalTurnover")) %>%
  group_by(Team, Statistic) %>%
  summarise(Total_Turnovers = sum(Total), .groups = "drop")

# Display the turnover breakdown 
View(turnover_breakdown)

# Create a table showing turnover data 
datatable(turnover_data)

# I initially created a bar chart that simply gave me just the total turnovers but I wanted something to break this down further 
# I researched and decided to creat a Stacked Bar Chart that show the amount of turnovers and breakdown of these turnovers
# Shows which type of turnover (bad passes, bad hands, missed goal turnovers) is most common per team.
# This would help coaches identify specific areas of improvement in realtion to turnovers 

```

Column {data-width=350}
-----------------------------------------------------------------------

### Chart H
ggplot(turnover_breakdown, aes(x = reorder(Team, -Total_Turnovers), y = Total_Turnovers, fill = Statistic)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  labs(title = "Turnover Breakdown by Type per Team", x = "Team", y = "Total Turnovers") +
  theme_minimal() +
  scale_fill_viridis_d(option = "C")

```{r}
# ---------------------------------------------------------------------------- #
# 2b. Turnovers - Create a boxplot for turnover distribution per team
# ---------------------------------------------------------------------------- #

# This would provide a statistcal summary for coaches and which teams have the widest range of turnovers and outliers 
# It should show team with highest turnovers and outliers struggle the most
# Team turnovers vs team performance 


```

### Chart I
ggplot(turnover_breakdown, aes(x = Team, y = Total_Turnovers, fill = Team)) +
  geom_boxplot(alpha = 0.7, outlier.color = "red") +
  coord_flip() +
  labs(title = "Turnovers Distribution per Team", x = "Team", y = "Total Turnovers") +
  theme_minimal() +
  scale_fill_viridis_d(option = "C")

```{r}
# ---------------------------------------------------------------------------- #
# 3. Defensive strengths - Look at defensive strengths of the super netball teams
# ---------------------------------------------------------------------------- #
# Provide a sumarry of defensive statistics looking at intercepts, deflections, blocks, rebounds
# Define Defensive Statistics (Fix Missing Variable)
defensive_stats <- c("intercepts", "deflections", "blocks", "rebounds")

# Filter defensive statistics
defensive_data <- SuperNetball_data %>%
  filter(Statistic %in% defensive_stats) %>%
  group_by(Athlete) %>%
  summarise(Total_Defensive_Actions = sum(Total)) %>%
  arrange(desc(Total_Defensive_Actions))

# View defensive data (Check if it's correctly created)
head(defensive_data)


# Display full list of defender in ranking order from the dataset
view(defensive_data)

# Filter defensive statistics
defensive_data <- SuperNetball_data %>%
  filter(Statistic %in% defensive_stats)

# Summarize defensive performance by athlete
defensive_summary <- defensive_data %>%
  group_by(Athlete, Statistic) %>%
  summarise(Total_Defensive_Actions = sum(Total), .groups = "drop")

# Display defensive summary 
View(defensive_summary)

# ---------------------------------------------------------------------------- #
# Create a whisker plot which will show the spread of defensive actions across various statistics 
# ---------------------------------------------------------------------------- #
# I will highlight the outliers either exceptional or poor performances
# I will then compare the intercetps, deflections, blocks and rebounds within the plot
```

Column {data-width=650}
-----------------------------------------------------------------------

### Chart J
ggplot(defensive_summary, aes(x = Statistic, y = Total_Defensive_Actions, fill = Statistic)) +
  geom_boxplot(alpha = 0.7, outlier.color = "red", outlier.shape = 16) +
  scale_fill_viridis_d(option = "C") +
  labs(title = "Distribution of Defensive Actions Across Players",
       x = "Defensive Statistic",
       y = "Total Defensive Actions") +
  theme_classic() +
  theme(
    axis.text.x = element_text(size = 12, angle = 45, hjust = 1, face = "bold"),
    axis.text.y = element_text(size = 12, face = "bold"),
    axis.title = element_text(size = 14, face = "bold"),
    plot.title = element_text(size = 16, face = "bold", hjust = 0.5)
  )

```{r}
# ---------------------------------------------------------------------------- #
# Possession Play - focus on midcourt performance and passing efficency, providing a summary of passing and feeds 
# ---------------------------------------------------------------------------- #
# Possession of the Netball is the key to the sport and key defenders are paramount to the teams success by gaining possession.  Coaches want to see their teams with the Netball so crucial area for coaches 

# Define midcourt-related statistics
midcourt_stats <- c("feeds", "feedWithAttempt", "centrePassReceives")

# Filter and summarize midcourt data
midcourt_data <- SuperNetball_data %>%
  filter(Statistic %in% midcourt_stats) %>%
  group_by(Athlete) %>%
  summarise(Total_Feeds = sum(Total, na.rm = TRUE)) %>%
  arrange(desc(Total_Feeds))

# View the entire mid court data set
view(midcourt_data)

# Display Top 10 Midcourt Players
head(midcourt_data, 10)

# Create a visualisation about the influence the midcourt has on the super netball league 
# Dispaly which players are creatig goal scoring opportunities 

```

Column {data-width=350}
-----------------------------------------------------------------------

### Chart K
ggplot(head(midcourt_data, 10), aes(x = reorder(Athlete, -Total_Feeds), y = Total_Feeds, fill = Athlete)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  labs(title = "Top 10 Playmakers (Feeds)", x = "Athlete", y = "Total Feeds") +
  theme_minimal()

```{r}
# Define Midcourt-Related Statistics
midcourt_stats <- c("feeds", "feedWithAttempt", "centrePassReceives")

# Filter the data to focus on midcourt statistics
midcourt_data <- SuperNetball_data %>%
  filter(Statistic %in% midcourt_stats)

# Summarize Midcourt Performance by Team
midcourt_team_summary <- SuperNetball_data %>%
  filter(Statistic %in% midcourt_stats) %>%
  group_by(Team, Statistic) %>%
  summarise(Total_Actions = sum(Total), .groups = "drop")

# Check the summary table
head(midcourt_team_summary)

# Check if the filtering was successful
head(midcourt_data)

# Create grouped bar chart with a summary of midcourt statistics providing a team comparison with easy side by side comparison with feeds, feed attempts and centre pass receives
# Coaches can identify which teams have strong midcourt playmakers and which aspects they are very good.
# Centre pass receives indicate strong ball progression from the centre
# Higher feed attempts may suggest aggressive attacking style of  play
```

### Chart L
ggplot(midcourt_team_summary, aes(x = Team, y = Total_Actions, fill = Statistic)) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_fill_viridis_d(option = "C") +
  labs(title = "Midcourt Performance by Team: Feeds, Feed Attempts & Centre Pass Receives",
       x = "Team",
       y = "Total Actions",
       fill = "Midcourt Statistic") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(size = 12, angle = 45, hjust = 1, face = "bold"),
    axis.text.y = element_text(size = 12, face = "bold"),
    axis.title = element_text(size = 14, face = "bold"),
    plot.title = element_text(size = 16, face = "bold", hjust = 0.5)
  )

```{r}
# ---------------------------------------------------------------------------- #
# 4. Fouls and Discipline - create a summary and visualisation of penalties by a player
# ---------------------------------------------------------------------------- #
# Summarise penalties by player
penalty_data <- SuperNetball_data %>%
  filter(Statistic %in% c("contactPenalties", "obstructionPenalties", "offsides", "breaks")) %>%
  group_by(Athlete, Statistic) %>%
  summarise(Total_Penalties = sum(Total), .groups = "drop") %>%
  arrange(desc(Total_Penalties))

# Check if data contains the correct columns
head(penalty_data)

# Display the penalised player data 
view(penalty_data)

# Display the top 10 penalised players 
head(penalty_data, 10)

penalty_data <- SuperNetball_data %>%
  filter(Statistic %in% c("contactPenalties", "obstructionPenalties", "offsides", "breaks")) %>%
  group_by(Athlete, Statistic) %>%
  summarise(Total_Penalties = sum(Total), .groups = "drop") %>%
  arrange(desc(Total_Penalties)) %>%
  slice_head(n = 10)

ggplot(penalty_data, aes(x = reorder(Athlete, -Total_Penalties), y = Total_Penalties, fill = Statistic)) +
  geom_bar(stat = "identity", width = 0.8) +
  coord_flip() +
  scale_fill_viridis_d(option = "C") +
  labs(title = "Top 10 Most Penalized Players", x = "Athlete", y = "Total Penalties") +
  theme_minimal()

# Create a tabke with top 10 penalised players 
datatable(penalty_data)

# Create a visualisation of the players with the most penalities 
# Visualization: Players with Most Penalties

# Create Bar Chart
```

### Chart M
ggplot(head(penalty_data, 10), aes(x = reorder(Athlete, -Total_Penalties), y = Total_Penalties, fill = Athlete)) +
  geom_bar(stat = "identity", width = 0.8) +  # Adjust width for better spacing
  coord_flip() +
  scale_fill_viridis_d(option = "C", begin = 0.2, end = 0.9) +  # Colorblind-friendly colors
  labs(title = "Top 10 Most Penalized Players",
       x = "Athlete",
       y = "Total Penalties",
       fill = "Athlete") +  # Legend for athletes
  theme_minimal() +
  theme(
    axis.text.x = element_text(size = 12, face = "bold"),
    axis.text.y = element_text(size = 12, face = "bold"),
    axis.title = element_text(size = 14, face = "bold"),
    plot.title = element_text(size = 16, face = "bold", hjust = 0.5),
    legend.position = "none"  # Hides legend since names are already on the y-axis
  )

```{r}
#------------------------------------------------------------------------------#
# 5. Discipline - Find a way to display discipline related penalities and fouls
#------------------------------------------------------------------------------#
# Filter penalty related statistics, obstrcution penalities, offsides, breaks
penalty_stats <- c("contactPenalties", "obstructionPenalties", "offsides", "breaks")

# Check the structure of the dataset
head(penalty_data)

# Create a whisker plot which will show variation in penalty counts and help to identify which penalty types are high or vary. It show outliers and highlights players that commit unusally high number of specific foul.
# This shows contact and obstruction have higher median, meaning more frequent of the penalitie.  Can allow coaches to focus on specific infractions and discipline areas

```

### Chart N
ggplot(penalty_data, aes(x = Statistic, y = Total_Penalties, fill = Statistic)) +
  geom_boxplot(alpha = 0.7, outlier.color = "red", outlier.shape = 16) +
  scale_fill_viridis_d(option = "C") +
  labs(title = "Whisker Plot: Distribution of Penalties Across Teams",
       x = "Penalty Type",
       y = "Total Penalties") +
  theme_classic() +
  theme(
    axis.text.x = element_text(size = 12, angle = 45, hjust = 1, face = "bold"),
    axis.text.y = element_text(size = 12, face = "bold"),
    axis.title = element_text(size = 14, face = "bold"),
    plot.title = element_text(size = 16, face = "bold", hjust = 0.5)
  )

```{r}
